Coherent optical creation of a single molecule -- Supplemental material

### 2020-12-10 paper.tex

The molecular lifetime of $0.199(9)~\mathrm{ms}$ in Fig.~\ref{f-lifetime}a is measured by preparing the molecule with a $\pi$-pulse,
followed after a variable delay by a second dissociating $\pi$-pulse.
This result is
consistent with the decay of the Rabi oscillation in Fig.~\ref{f-raman}b. Preliminary experiments and theoretical considerations indicate that the molecular lifetime may be limited by two-photon coupling to the atomic continuum~\cite{YichaoYu}.  Atom loss is shown to be small in Fig.~\ref{f-lifetime}b by measuring the two-body lifetimes of the atoms directly without the second Raman frequency. In principle, destructive interference that reduces the Raman Rabi rate $\Omega_{\rm R}$ for negative detunings $\Delta$ changes to constructive interference for positive detunings, but additional molecular resonances make the positive region unusable.  More negative detunings that might reduce the scattering rate were prevented by vanishing of $\Omega_{\rm R}$.

>>>>

The molecular lifetime measurement in Fig.~\ref{f-lifetime}a is performed
by preparing the molecule with a $\pi$-pulse,
followed after a variable delay by a second dissociating $\pi$-pulse.
Together with the Rabi oscillation in Fig.~\ref{f-raman}b,
we obtain a molecule lifetime of $0.199(9)~\mathrm{ms}$.
Preliminary experiments and theoretical considerations indicate that
the molecular lifetime may be limited by two-photon coupling
to the atomic continuum~\cite{YichaoYu}.
Atom loss is shown to be small in Fig.~\ref{f-lifetime}b
by measuring the two-body lifetimes of the atoms directly
without the second Raman frequency.
In principle, destructive interference that reduces the Raman Rabi rate $\Omega_{\rm R}$
for negative detunings $\Delta$ changes to constructive interference
for positive detunings, but additional molecular resonances make the positive region unusable.
More negative detunings that might reduce the scattering rate
were prevented by vanishing of $\Omega_{\rm R}$.

### 2020-12-10 sm.tex

\section{Excited Molecular States} \label{sm:excited_states}
To obtain a full calculation for the expected Raman Rabi frequency, light shift and scattering, we need to account for all the excited molecular states.
The lowest eight excited molecular potentials in NaCs asymptote to the dissociation limit of Na 3S + Cs 6P.
These eight potentials can be written in the Hund's case (a) basis as $ \mathrm{c^3\Sigma^+_{\Omega = 0^-},c^3\Sigma^+_{\Omega = 1}, b^3\Pi_{\Omega = 0^-}, b^3\Pi_{\Omega = 0^+}, b^3\Pi_{\Omega = 1}, b^3\Pi_{\Omega = 2}, B^1\Pi_{\Omega = 1}, A^1\Sigma^+_{\Omega = 0^+}}$ \cite{Korek2007}.
Due to spin-orbit coupling, especially large in Cs, the excited molecular states are mixed, containing components in multiple potentials.
Fortunately, single-channel potential energy curves for $\mathrm{c^3\Sigma^+_1}$ and $  B^1\Pi_1 $ have been experimentally determined, and are predictive across the entire energy range of the potential \cite{Grochola2010, Grochola2011, Liu2017}.
Since there is only one $ \Omega = 2$ potential at this dissociation limit, a single-channel potential has been determined for $ b^3\Pi_2$ as well \cite{Zabawa2012}.
Unfortunately, single-channel potentials for the other electronic states do not exist.
A deperturbation analysis was required for states belonging to the mixed $\mathrm{A-b}$ complex,
and the fitted potentials and spin-orbit coupling coefficients are only accurate within the energy range of the measured experimental data \cite{Zaharova2009}.
Since our experiment attempts to work in a regime where the predominant Raman coupling comes from the $v' = 0$ state of $ \mathrm{c^3\Sigma^+_1} $,
the contribution of the other seven potentials is mainly to a background that is nearly detuning independent.
Thus, for computational ease, we use the diabatic potentials for $ \mathrm{b^3\Pi_{0}, b^3\Pi_{1}} $ and $ \mathrm{A^1\Sigma^+_{0^+}}$, and remove the states nearest to the $ v' = 0$ state of $ \mathrm{c^3\Sigma^+_1} $.
Lastly, to our knowledge, no potential fit to experimental data exists for the $ c^3\Sigma^+_{0^-}$ state. Assuming that its contribution is only to the background, as an approximation, we simply double the contribution of all states except $v' = 0$ from $ c^3\Sigma^+_1$.

\section{Calculating Matrix Elements with Coupled Channel Ground State}
To calculate Rabi frequencies for transitions to the eight excited molecular potentials, we
obtain their wavefunctions from a single-channel calculation \cite{Fattal1996} using the potential energy curves described in section \ref{sm:excited_states}. We
then calculate overlap integrals between these functions and either the triplet ($S=1$) or singlet ($S=0$) components of the
coupled-channel wavefunctions for either the target molecular state or the confined atomic state.
These calculations assume that the intermediate states have pure singlet or triplet character and neglect its
nuclear spin character, which is unknown. These calculations include theoretical electronic transition dipole moments\footnote{Private communication with R. Gonzalez-Ferez and R. Moszynski}

@@

In this calculation, there is a matrix element, $\Omega_{ai}, \Omega_{mi}$ between each channel $ i $ of the coupled channel ground state and an excited state. The total Raman Rabi frequency $ \Omega_{\rm R}$, light shift $ \delta$ and off-resonant scattering $ \Gamma_{\rm s} $ contribution from this single excited state is calculated as

\begin{align}
  \Omega_{\rm R} &= \displaystyle\sum_{i} \frac{\Omega_{ai} \Omega_{mi}}{2\Delta} \\
  \delta &= \displaystyle\sum_{i} \frac{\Omega_{ai}^2 - \Omega_{mi}^2}{2\Delta} \\
  \Gamma_{\rm s} &= \displaystyle\sum_{i} \Gamma_{\rm e} \frac{\Omega_{ai}^2 + \Omega_{mi}^2}{2\Delta^2}
\end{align}

@@

Since our coupled-channel calculations include the harmonic confinement, we can calculate the enhancement in $\Omega_{\rm a}$ with increasing confinement. Using two calculations with $ \omega_{\text{trap}} = 36.5~\mathrm{kHz}$ and $\omega_{\text{trap}} = 80~\mathrm{kHz}$ trapping frequencies, and assuming a power law scaling, we find $ \Omega_{\rm a} \propto \omega_{\text{trap}}^{0.58} $. Thus, $\Omega_{\rm a} \propto P^{0.29} $, which was also experimentally verified. However, we note that in the actual experiment, our system does not have spherically symmetric confinement. The radial trapping frequency is about $5.7$ times larger than the axial trapping frequency.
\todo{
  Power/intensity calibration
}

>>>>

\section{Excited Molecular States} \label{sm:excited_states}
To obtain a full calculation for the expected Raman Rabi frequency, light shift and scattering, we need to account for all the excited molecular states.
The lowest eight excited molecular potentials in NaCs asymptote to the dissociation limit of Na 3S + Cs 6P.
These eight potentials can be written in the Hund's case (a) basis as $ \mathrm{c^3\Sigma^+_{\Omega = 0^-},c^3\Sigma^+_{\Omega = 1}, b^3\Pi_{\Omega = 0^-}, b^3\Pi_{\Omega = 0^+}, b^3\Pi_{\Omega = 1}, b^3\Pi_{\Omega = 2}, B^1\Pi_{\Omega = 1}, A^1\Sigma^+_{\Omega = 0^+}}$ \cite{Korek2007}.
Due to spin-orbit coupling, especially large in Cs, the excited molecular states are mixed, containing components in multiple potentials.
Fortunately, single-channel potential energy curves for $\mathrm{c^3\Sigma^+_1}$ and $  B^1\Pi_1 $ have been experimentally determined, and are predictive across the entire energy range of the potential \cite{Grochola2010, Grochola2011, Liu2017}.
Since there is only one $ \Omega = 2$ potential at this dissociation limit, a single-channel potential has been determined for $ b^3\Pi_2$ as well \cite{Zabawa2012}.
Unfortunately, single-channel potentials for the other electronic states do not exist.
A deperturbation analysis was required for states belonging to the mixed $\mathrm{A-b}$ complex,
and the fitted potentials and spin-orbit coupling coefficients are only accurate within the energy range of the measured experimental data \cite{Zaharova2009}.
Since our experiment attempts to work in a regime where the predominant Raman coupling comes from the $v' = 0$ state of $ \mathrm{c^3\Sigma^+_1} $,
the contribution of the other seven potentials is mainly to a background that is nearly detuning independent.
Thus, for computational ease, we use the diabatic potentials for $ \mathrm{b^3\Pi_{0}, b^3\Pi_{1}} $ and $ \mathrm{A^1\Sigma^+_{0^+}}$.
%and remove the states nearest to the $ v' = 0$ state of $ \mathrm{c^3\Sigma^+_1} $.
Lastly, to our knowledge, no potential fit to experimental data exists for the $ c^3\Sigma^+_{0^-}$ state. Assuming that its contribution is only to the background, as an approximation, we simply double the contribution of all states except $v' = 0$ from $ c^3\Sigma^+_1$.

\section{Calculating Matrix Elements with Coupled Channel Ground State}
To calculate Rabi frequencies for transitions to the eight excited molecular potentials, we
obtain their wavefunctions from a single-channel calculation \cite{Fattal1996} using the potential energy curves described in section \ref{sm:excited_states}. We
then calculate overlap integrals between these functions and either the triplet ($S=1$) or singlet ($S=0$) components of the
coupled-channel wavefunctions for either the target molecular state or the confined atomic state.
These calculations assume that the intermediate states have pure singlet or triplet character and neglect its
nuclear spin character, which is unknown. These calculations include theoretical electronic transition dipole moments~\footnote{Private communication with R. Gonzalez-Ferez and R. Moszynski}.

@@

In this calculation, there is a matrix element, $\Omega_{\mathrm{a}i}, \Omega_{\mathrm{m}i}$ between each channel $ i $ of the coupled channel ground state and an excited state. The total Raman Rabi frequency $ \Omega_{\rm R}$, light shift $ \delta$ and off-resonant scattering $ \Gamma_{\rm s} $ contribution from this single excited state is calculated as

\begin{align}
  \Omega_{\rm R} &= \displaystyle\sum_{i} \frac{\Omega_{\mathrm{a}i} \Omega_{\mathrm{m}i}}{2\Delta} \\
  \delta &= \displaystyle\sum_{i} \frac{\Omega_{\mathrm{a}i}^2 - \Omega_{\mathrm{m}i}^2}{2\Delta} \\
  \Gamma_{\rm s} &= \displaystyle\sum_{i} \Gamma_{\rm e} \frac{\Omega_{\mathrm{a}i}^2 + \Omega_{\mathrm{m}i}^2}{2\Delta^2}
\end{align}

@@

Since our coupled-channel calculations include the harmonic confinement, we can calculate the enhancement in $\Omega_{\rm a}$ with increasing confinement. Using two calculations with $ \omega_{\text{trap}} = 36.5~\mathrm{kHz}$ and $\omega_{\text{trap}} = 80~\mathrm{kHz}$ trapping frequencies, and assuming a power law scaling, we find that the wavefunction overlap scales as $ \omega_{\text{trap}}^{0.58} = P^{0.29} $. Thus, $\Omega_{\rm a} \propto P^{0.79} $, which was also experimentally verified. However, we note that in the actual experiment, our system does not have spherically symmetric confinement. The radial trapping frequency is about $5.7$ times larger than the axial trapping frequency.
\todo{
  Power/intensity calibration
}
